Doppler Imaging Fun

### 2019-06-10 tex/ms.tex
woot
typo
figures, typos, etc

If we Taylor expand Equation~(\ref{eq:IntensityUnif}) about $\beta = 0$, we obtain
%
\begin{align}
    \label{eq:TaylorUnifExplicit}
    I(\lambda, x, y) 
        &=
        I_0(\lambda_0) \Bigg|_{\beta=0}
        + 
        \frac{\mathrm{d}I_0(\lambda_0)}{\mathrm{d}\beta} \Bigg|_{\beta=0} \Delta\beta(x, y)
        + 
        \frac{\mathrm{d}^2I_0(\lambda_0)}{\mathrm{d}\beta^2} \Bigg|_{\beta=0} \Delta\beta(x, y)^2
        +
        ... 
\end{align}
%
The derivatives of the spectrum $I_0(\lambda_0)$ with respect to
$\beta$ are found by application of Fa\`a di Bruno's formula for taking high
order derivatives of the chain rule:
%
\begin{proof}{faa_di_bruno}
    \label{eq:dIdbeta}
    \frac{\mathrm{d}^n I_0(\lambda_0)}{\mathrm{d}\beta^n} \Bigg|_{\beta=0} &=
    \sum_{k=1}^n \frac{\mathrm{d}^k I_0(\lambda_0)}{\mathrm{d}\lambda_0^k} \lambda_0^k P_{nk}
\end{proof}
%
where
%
\begin{proof}{faa_di_bruno}
    \label{eq:Pnk}
    P_{nk} \equiv B_{n, k}\Bigg( \Big\{(-1)^j j! \Big\}_{j=1}^{n - k + 1} \Bigg)
\end{proof}
%
is an integer computed from the incomplete Bell polynomial $B_{n, k}$.
Given this result, and noting that $\lambda_0 = \lambda$ when $\beta = 0$,
we may re-write Equation~(\ref{eq:TaylorUnifExplicit}) as
%
\begin{align}
    \label{eq:TaylorUnifSum}
    I(\lambda, x, y) 
        &=
        I_0(\lambda)
        + 
        \sum_{n=1}^\infty
            \Bigg(
                \sum_{k=1}^n \frac{\mathrm{d}^k I_0(\lambda)}{\mathrm{d}\lambda^k} \lambda^k P_{nk} 
            \Bigg)
            \Bigg(
                \Delta\beta(x, y)^n
            \Bigg)
\end{align}
%
Assuming we only have access to spatially unresolved spectra of the star, the quantity
we observe is the integral of the intensity over the surface of the star, which we
denote $S$:
%
\begin{align}
    \label{eq:TaylorUnifSumIntegral}
    S(\lambda) 
        &=
        \pi I_0(\lambda)
        + 
        \sum_{n=1}^\infty
            \Bigg(
                \sum_{k=1}^n \frac{\mathrm{d}^k I_0(\lambda)}{\mathrm{d}\lambda^k} \lambda^k P_{nk} 
            \Bigg)
            \Bigg(
                \int{\Delta\beta(x, y)^n}\mathrm{d}\Omega
            \Bigg)
\end{align}
%
where $\int\mathrm{d}\Omega$ denotes the surface integral over the visible disk
of the star.

@@

As an example of an application of this equation, consider the case of a rigidly rotating,
uniform star with equatorial velocity in units of the speed of light equal to $\beta_\mathrm{eq}$. 
It is easy to show that in this case, $\beta(x, y) = \beta_\mathrm{eq} x$. The integrals
in Equation~(\ref{eq:TaylorUnifSumIntegral}) are trivial to compute:
%
\begin{align}
    \int{\Delta\beta(x, y)^n}\mathrm{d}\Omega = 
        \begin{cases} 
            0 & \quad\quad\quad\quad\quad n \, \mathrm{odd} \\
            \beta_\mathrm{eq}\frac{\Gamma\big(\frac{n}{2} + \frac{1}{2}\big)}{\Gamma\big(\frac{n}{2} + 2\big)} & \quad\quad\quad\quad\quad n \, \mathrm{even}
        \end{cases}
\end{align}
%
so the expression for the spectrum becomes
\begin{align}
    \label{eq:LineBroadenedSpectrum}
    S(\lambda) 
        &=
        \pi I_0(\lambda)
        + 
        \beta_\mathrm{eq}
        \sum_{n\,\mathrm{even}}
            \frac{\Gamma\big(\frac{n}{2} + \frac{1}{2}\big)}{\Gamma\big(\frac{n}{2} + 2\big)}
            \sum_{k=1}^n 
                \frac{\mathrm{d}^k I_0(\lambda)}{\mathrm{d}\lambda^k} \lambda^k P_{nk} 
\end{align}
%
Equation~(\ref{eq:LineBroadenedSpectrum}) is the analytic expression for line broadening
due to a rotating star.

>>>>

If we Taylor expand Equation~(\ref{eq:IntensityUnif}) about $\beta = 0$, we obtain
%
\begin{align}
    \label{eq:TaylorUnifExplicit}
    I(\lambda, x, y) 
        &=
        I_0(\lambda_0) \Bigg|_{\beta=0}
        + 
        \frac{\mathrm{d}I_0(\lambda_0)}{\mathrm{d}\beta} \Bigg|_{\beta=0} \Delta\beta(x, y)
        + 
        \frac{1}{2}\frac{\mathrm{d}^2I_0(\lambda_0)}{\mathrm{d}\beta^2} \Bigg|_{\beta=0} \Delta\beta(x, y)^2
        +
        ... 
\end{align}
%
The derivatives of the spectrum $I_0(\lambda_0)$ with respect to
$\beta$ are found by application of Fa\`a di Bruno's formula for taking high
order derivatives of the chain rule:
%
\begin{proof}{faa_di_bruno}
    \label{eq:dIdbeta}
    \frac{\mathrm{d}^n I_0(\lambda_0)}{\mathrm{d}\beta^n} \Bigg|_{\beta=0} &=
    \sum_{k=1}^n P_{nk} \frac{\mathrm{d}^k I_0(\lambda_0)}{\mathrm{d}\lambda_0^k} \lambda_0^k
\end{proof}
%
where
%
\begin{proof}{faa_di_bruno}
    \label{eq:Pnk}
    P_{nk} \equiv B_{n, k}\Bigg( \Big\{(-1)^j j! \Big\}_{j=1}^{n - k + 1} \Bigg)
\end{proof}
%
is an integer computed from the incomplete Bell polynomial $B_{n, k}$.
Given this result, and noting that $\lambda_0 = \lambda$ when $\beta = 0$,
we may re-write Equation~(\ref{eq:TaylorUnifExplicit}) as
%
\begin{align}
    \label{eq:TaylorUnifSum}
    I(\lambda, x, y) 
        &=
        I_0(\lambda)
        + 
        \sum_{n=1}^\infty
            \Bigg(
                \sum_{k=1}^n \frac{P_{nk}}{n!} \frac{\mathrm{d}^k I_0(\lambda)}{\mathrm{d}\lambda^k} \lambda^k
            \Bigg)
            \Bigg(
                \Delta\beta(x, y)^n
            \Bigg)
\end{align}
%
Assuming we only have access to spatially unresolved spectra of the star, the quantity
we observe is the integral of the intensity over the surface of the star, which we
denote $S$:
%
\begin{align}
    \label{eq:TaylorUnifSumIntegral}
    S(\lambda) 
        &=
        \pi I_0(\lambda)
        + 
        \sum_{n=1}^\infty
            \Bigg(
                \sum_{k=1}^n \frac{P_{nk}}{n!} \frac{\mathrm{d}^k I_0(\lambda)}{\mathrm{d}\lambda^k} \lambda^k
            \Bigg)
            \Bigg(
                \int{\Delta\beta(x, y)^n}\mathrm{d}\Omega
            \Bigg)
\end{align}
%
where $\int\mathrm{d}\Omega$ denotes the surface integral over the visible disk
of the star.

@@

As an example of an application of this equation, consider the case of a rigidly rotating,
uniform star with equatorial velocity in units of the speed of light equal to $\beta_\mathrm{eq}$. 
It is easy to show that in this case, $\beta(x, y) = \beta_\mathrm{eq} x$. The integrals
in Equation~(\ref{eq:TaylorUnifSumIntegral}) are straightforward to compute:
%
\begin{align}
    \int{\Delta\beta(x, y)^n}\mathrm{d}\Omega = 
        \begin{cases} 
            0 & \quad\quad\quad\quad\quad n \, \mathrm{odd} \\
            \beta_\mathrm{eq}^n\sqrt{\pi}\frac{\Gamma\big(\frac{n}{2} + \frac{1}{2}\big)}{\Gamma\big(\frac{n}{2} + 2\big)} & \quad\quad\quad\quad\quad n \, \mathrm{even}
        \end{cases}
\end{align}
%
so the expression for the spectrum becomes
\begin{align}
    \label{eq:LineBroadenedSpectrum}
    S(\lambda) 
        &=
        \pi I_0(\lambda)
        + 
        \sqrt{\pi}
        \sum_{n\,\mathrm{even}>0}
            \beta_\mathrm{eq}^n
            \frac{\Gamma\big(\frac{n}{2} + \frac{1}{2}\big)}{\Gamma\big(\frac{n}{2} + 2\big)}
            \sum_{k=1}^n 
            \frac{P_{nk}}{n!} \frac{\mathrm{d}^k I_0(\lambda)}{\mathrm{d}\lambda^k} \lambda^k 
\end{align}
%
Equation~(\ref{eq:LineBroadenedSpectrum}) is the analytic expression for line broadening
due to a uniform rotating star.

\begin{figure}[p!]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/convergence.pdf}
    \oscaption{convergence}{%
        Fractional root mean square error on the Taylor expansion of a
        Doppler-shifted Gaussian absorption line profile as a function of 
        the Doppler parameter 
        $\xi \equiv \beta / \left( \nicefrac{\sigma}{\mu}\right)$ 
        and the order of the expansion $N$. The Taylor series
        does not converge above $\xi \approx 3$.
        \label{fig:convergence}
    }
\end{figure}

\begin{figure}[p!]
    \centering
    \subfigure[Local absorption]{\label{fig:convergence_line:a}%
        \includegraphics[width=2.5in]{figures/local_spectrum.pdf}}
    \hspace{0.25in}
    \subfigure[Disk-integrated absorption]{\label{fig:convergence_line:b}%
        \includegraphics[width=2.5in]{figures/disk_integrated_spectrum.pdf}}
    \oscaption{convergence}{%
        Taylor expansion of a Doppler-shifted Gaussian absorption line
        (orange) for different values of the Doppler parameter 
        $\xi \equiv \beta / \left( \nicefrac{\sigma}{\mu}\right)$. 
        The dashed black curve is the line in the rest frame, and the blue
        curve is the shifted line computed numerically at high precision.
        \label{fig:convergence_line}
    }
\end{figure}
